---
title: "README"
author: "Aaron Cole"
date: "21 December 2015"
output:
    md_document:
        variant: markdown_github
---

# README

## Project information
* Coursera - Data Science Specialization.
* Course: Getting and Cleaning Data.
* Getting and Cleaning Data Course Project.
* Script author: Aaron Cole.
* Script date: 21-Dec-2015.


## Preamble
This project prepares and tidies the [Activity Recognition Using Smartphones Dataset](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones#). Call the **run_analysis()** function from the **run_analysis.R** file to prepare and tidy the data. Once tidy, the data is summarised by the mean for each measurement, for each activity, for each student and then output as the file **tidy_data.CSV**. The generated summary data can also be returned to the caller as a dplyr tbl_df, by passing the first argument as TRUE when making the function call. Second and third logical arguments can be set for re-preparing and re-tidying the data.

## Libraries
The following libraries have been used. They are loaded at the top off the **run_analysis.R** file.
```{r eval=FALSE}
# Load libraries.
library(dplyr, quietly = TRUE, verbose = FALSE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, verbose = FALSE, warn.conflicts = FALSE)
```

## Scripts summary
Script | Summary
-------|--------
run_analysis.R | This is the main entry point script. It makes calls to the prepare and tidy scripts. Can be called with arguments to return the tidy data, re-prepare the data, or re-tidy the data. This script is responsible for creating the local files of the prepared and tidied data.
prepare_data.R | Takes an argument value of "test" or "train" to indicate which dataset is to be prepared and returned.
tidy_data.R | Takes an argument value of the prepared data, tidies it, then summarises it for return.

## Script annotations
All three scripts have been comprehensively annotated to describe their functionality.

## CodeBook
The CodeBook.md file provides a descriptive outline of what the scripts are doing. Explanation of the data and the transformations that have been carried out is also provided.

## Usage
The following call to the **run_analysis** function will re-prepare and re-tidy the data before returning it. The call has been wrapped within a call to str() for presentation purposes.
```{r eval=TRUE, warning=FALSE}
source("run_analysis.R")
str(run_analysis(TRUE, FALSE, FALSE))
```

## Output
* As per the example above, the data that can be returned from **run\_analysis** is a dplyr tbl_df, with 180 observations and 68 variables.
* The "subject" variable is type "integer".
* The "activity" variable is type "factor", with 6 levels.
* The remaining 66 variables are all type "numeric".
* More information about the returned tidy summary data can be found in **CookBook.md**.