---
title: "README"
author: "Aaron Cole"
date: "20 December 2015"
output: 
    html_document:
        toc: true
        theme: united
---

## Project information
* Coursera - Data Science Specialization
* Course: Getting and Cleaning Data
* Getting and Cleaning Data Course Project
* Script author: Aaron Cole
* Script date: 20-Dec-2015


## Preamble
This project prepares and tidies the "Activity Recognition Using Smartphones Dataset". Once tidy, the data is summarised and then output as **tidy_data.CSV** as well as being returned from the function as a dplyr tbl_df.

## Libraries
The following libraries have been used. They are loaded at the top off the **run_analysis.R** file.
```{r eval=FALSE}
    # Source scripts.
    source("prepare_data.R")
    source("tidy_data.R")
```

## Script summary
Script | Summary
-------|--------
run_analysis.R | Main script, which makes calls to the prepare and tidy scripts. Can be called with arguments to return the tidy data, re-prepare the data, or re-tidy the data. This script is responsible for creating the local files of the prepared and tidied data.
prepare_data.R | Takes an argument value of "test" or "train" to indicate which dataset is to be prepared and returned.
tidy_data.R | Takes the prepared dataset, tidies it, then summarises it for return.

## Script annotations
All three scripts have been comprehensively annotated to describe their functionality.

## CodeBook
The CodeBook.Rmd file provides a descriptive outline of what the scripts are doing. Explanation of the data and the transformations that have been carried out is also provided.

## Usage
The following call to the **run_analysis** function will re-prepare and re-tidy the data before returning it.
```{r warning=FALSE}
source("run_analysis.R")
run_analysis(TRUE, TRUE, TRUE)
```

